Nodal Discontinuous Galerkin Method for the Two-Moment Model of Neutrino Transport

### 2017-07-25 Documents/M1/dgM1Curvilinear.tex
Minor updates.
Added figures and text for the line source problem.  Also some text on the discontinuity detector.

The limiting procedure first computes, for each dimension $i=1,\ldots,d$, the limited slopes $\widetilde{\bcC}_{|\vect{1}|}^{i}$ by comparing the variation within $\bK$, $\bcC_{|\vect{1}|}^{i}$, with the variation in the cell average by considering the neighboring elements
\begin{equation}
  \widetilde{\bcC}_{|\vect{1}|}^{i}
  =\text{minmodB}
  \big(\,
    \bcC_{|\vect{1}|}^{i},\,
    \beta\,(\langle\bcM\rangle_{\bK}-\langle\bcM\rangle_{\bK}^{i,-}),\,
    \beta\,(\langle\bcM\rangle_{\bK}^{i,+}-\langle\bcM\rangle_{\bK})
  \,\big),
  \label{eq:limitedSlopes}
\end{equation}
where the limiter parameter $\beta\in[1,2]$ controls how aggressive limiting is applied.  
($\beta=2$ is the least aggressive, and, in many cases, leads to the most accurate results.)  
The $\text{minmodB}$ function is given by \citet{cockburnShu_2001} as
\begin{equation}
  \mbox{minmodB}(a,b,c)
  =\left\{\begin{array}{ll}
  a \quad\text{if}\quad |a|\le M\,\Delta x^{2} \\
  \mbox{minmod}(a,b,c) \quad \text{otherwise},
  \end{array}\right.
  \label{eq:minmodB}
\end{equation}
and the usual minmod function is given by
\begin{equation}
  \mbox{minmod}(a,b,c)
  =\left\{\begin{array}{ll}
  s\,\min[\,|a|,|b|,|c|\,] \quad\text{if}\quad s=\mbox{sign}(a)=\mbox{sign}(b)=\mbox{sign}(c) \\
  0 \quad \text{otherwise}.
  \end{array}\right.
  \label{eq:minmod}
\end{equation}
In Eq.~\eqref{eq:minmodB}, the parameter $M>0$ is proportional to the second derivative, and can be adjusted to prevent clipping at smooth extrema.  
For $M=0$, the $\text{minmodB}$ functions reduces to the usual minmod function in Eq.~\eqref{eq:minmod}.  
In Eq.~\eqref{eq:limitedSlopes}, $\langle\bcM\rangle_{\bK}^{i,-/+}$ is the cell average in the element to the left/right in the $i$-th dimension of the element under consideration.

@@

\ee{Discuss discontinuity detection using Harten's sub-cell resolution method.}

@@

As we mentioned in Section~\ref{sec:twoMomentModel} when discussing Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}, the total opacity (excluding NES) is $\kappa(\epsilonNu,\bU)=\tilde{\chi}(\epsilonNu,\bU)+\sigma(\epsilonNu,\bU)$, where the absorption opacity $\tilde{\chi}(\epsilonNu,\bU)$, corrected for stimulated absorption, is given by the sum of the neutrino emissivity $\eta(\epsilonNu,\bU)$ and the neutrino absorptivity $\chi(\epsilonNu,\bU)$
\begin{equation}
  \tilde{\chi}(\epsilonNu,\bU) = \eta(\epsilonNu,\bU) + \chi(\epsilonNu,\bU). 
\end{equation}
The scattering opacity $\sigma(\epsilonNu,\bU)$ is due to (isotropic) elastic scattering.   
Neutrino-electron scattering gives rise to the integral operators involving the rates $R^{\In}(\epsilonNu,\epsilonNu',\bU)$ and $R^{\In}(\epsilonNu,\epsilonNu',\bU)$.  
(In the following, to simplify the notation, we will drop the explicit spatial dependence and dependence on material properties $\bU$.)  
To arrive at the collision terms for the two-moment model, we consider the space-homogeneous Boltzmann equation \citep[e.g., Eq.~(7) in][]{mezzacappaBruenn_1993c}
\begin{align}
  \f{1}{c}\deriv{f}{t}(\omegaNu,\epsilonNu,t)
  & = \eta(\epsilonNu) - \tilde{\chi}(\epsilonNu)\,f(\omegaNu,\epsilonNu,t) \nonumber \\
  & \hspace{12pt}
  + \frac{\epsilonNu^2}{c\,(hc)^3} \int_{\bbS^{2}} R_{\IS}(\bl'\cdot\bl,\epsilonNu)\,f(\omega', \epsilonNu,t)\,d\omegaNu'
  - \frac{\epsilonNu^2}{c\,(hc)^3} f(\omegaNu,\epsilonNu,t) \int_{\bbS^{2}} R_{\IS}(\bl'\cdot\bl,\epsilonNu)\,d\omega' \nonumber \\
  & \hspace{12pt} 
  + \frac{1}{c\,(hc)^3}\big(1- f(\omegaNu,\epsilonNu)\big) \, 
  \int_{\bbR^{+}} \int_{\bbS^{2}}  R^{\In}_{\NES} (\bl'\cdot\bl,\epsilonNu,\epsilonNu')\,f(\omega',\epsilonNu',t)\,d\epsilonNu' \epsilonNu'^{2}\,d\omega' \nonumber \\
  & \hspace{12pt} 
  - \frac{1}{c\,(hc)^3} f(\omegaNu,\epsilonNu,t) \int_{\bbR^{+}} 
  \int_{\bbS^{2}}  R^{\Out}_{\NES} (\bl'\cdot\bl,\epsilonNu,\epsilonNu^{\prime})\,\big(1-f(\omegaNu',\epsilonNu',t )\big)\,d\epsilonNu' \epsilonNu'^{2}\,d\omegaNu',
  \label{eq:homogeneousBoltzmann}
\end{align}
where we have brought back physical constants; i.e., the speed of light $c$ and the Planck constant $h$.  
In the scattering kernels, $\bl(\omegaNu)=(\cos\vartheta,\sin\vartheta\cos\varphi,\sin\vartheta\sin\varphi)^{\mbox{\tiny T}}$, so that
\begin{equation}
 \bl'\cdot\bl  = \mu\mu' + [(1-\mu^2)(1-\mu'^{2})]^{1/2}\cos(\phiNu-\phiNu')
 \equiv \cos\alpha,
 \quad(\mu=\cos\thetaNu)
\end{equation}
is the cosine of the angle between incoming and outgoing neutrino is the scattering process.  
In Eq.~\eqref{eq:homogeneousBoltzmann}, the first two terms are due to emission and absorption, the third and fourth terms are due to elastic (isoenergetic) neutrino scattering on nucleons and nuclei, and the last two terms are due to inelastic (non-isoenergetic) neutrino-electron scattering.

@@

The NES scattering kernel $R_{\NES}^{\Out}(\cos\alpha,\epsilonNu,\epsilonNu')$ is given by Eq.~(9) in \citet{mezzacappaBruenn_1993c}.  
As we did in the isoenergetic scattering case, we approximate the angular dependence of the scattering kernels by a two-term Legendre expansion; e.g.,
\begin{equation}
  R^{\Out}_{\NES}(\cos\alpha,\epsilonNu,\epsilonNu') 
  \approx \f{1}{2} \Phi_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu') + \f{3}{2}\Phi^{\Out}_{\NES,1}(\epsilonNu,\epsilonNu')\,\cos\alpha,
\end{equation}
where
\begin{equation}
  \Phi_{\NES,\ell}^{\Out}(\epsilonNu,\epsilonNu')
  =\int_{-1}^{+1}R_{\NES}^{\Out}(\cos\alpha,\epsilonNu,\epsilonNu')\,P_{\ell}(\cos\alpha)\,d\cos\alpha \quad (\ell=0,1).  
\end{equation}
(Again, these integrals are computed numerically using a ??-point Gauss-Legendre quadrature.  (\ee{Ran: Fill in the number of quadrature points}.))
The in-scattering kernels $\Phi_{\NES,\ell}^{\In}(\epsilonNu,\epsilonNu')$ are obtained from symmetry properties \citep{cernohorsky_1994}
\begin{equation}
  \Phi_{\NES,\ell}^{\In}(\epsilonNu,\epsilonNu')=\Phi_{\NES,\ell}^{\Out}(\epsilonNu',\epsilonNu)
  \quad\text{and}\quad
  \Phi_{\NES,\ell}^{\Out}(\epsilonNu',\epsilonNu)
  =e^{-(\epsilonNu-\epsilonNu')/k_{\mbox{\tiny B}}T}\,\Phi_{\NES,\ell}^{\Out}(\epsilonNu,\epsilonNu').
  \label{eq:symmetryNES}
\end{equation}
Inserting the Legendre expansion into Eq.~\eqref{eq:homogeneousBoltzmann}, and taking the zeroth moment (number equation), gives
\begin{align}
  \f{1}{c}\deriv{\cJ}{t}
  &=\big(\,4\pi-\cJ(\epsilonNu)\,\big)\int_{\bbR^{+}}R_{\NES,0}^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  -\bcH(\epsilonNu)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\In}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'} \nonumber \\
  & \hspace{12pt}
  -\cJ(\epsilonNu)\int_{\bbR^{+}}R_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu')\,\big(\,4\pi-\cJ(\epsilonNu')\,\big)\,dV_{\epsilonNu'}
  +\bcH(\epsilonNu)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\Out}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'}.  
\end{align}
Similarly, taking the first moment (number flux) of Eq.~\eqref{eq:homogeneousBoltzmann} results in
\begin{align}
  \f{1}{c}\deriv{\bcH}{t}
  &=-\bcH(\epsilonNu)\int_{\bbR^{+}}R_{\NES,0}^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  +\big(\,(4\pi/3)\,\bI-\bcK\,\big)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\In}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'} \nonumber \\
  & \hspace{12pt}
  -\bcH\int_{\bbR^{+}}R_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu')\,\big(\,4\pi-\cJ(\epsilonNu')\,\big)\,dV_{\epsilonNu'}
  +\bcK\cdot\int_{\bbR^{+}}R_{\NES,1}^{\Out}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'},
\end{align}
where we have defined
\begin{equation}
  R_{\NES,0}^{\In/\Out}
  =\f{1}{c\,(hc)^{3}}\,\f{1}{2}\,\Phi_{\NES,0}^{\In/\Out}(\epsilonNu,\epsilonNu')
  \quad\text{and}\quad
  R_{\NES,1}^{\In/\Out}
  =\f{1}{c\,(hc)^{3}}\,\f{3}{2}\,\Phi_{\NES,1}^{\In/\Out}(\epsilonNu,\epsilonNu').  
\end{equation}
We tabulate the scattering kernels $R_{\NES,0}^{\Out}$ and $R_{\NES,1}^{\Out}$, which have units MeV$^{-3}$ cm$^{-1}$, and obtain the in-scattering rates from the symmetry properties in Eq.~\eqref{eq:symmetryNES}.  
We only use $R_{\NES,0}^{\In/\Out}$ in the numerical experiments in Section~\ref{sec:numericalTableOpacities}, so that $R^{\In/\Out}=R_{\NES,0}^{\In/\Out}$ in Eqs~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}.

@@

>>>>

The limiting procedure first computes, for each dimension $i=1,\ldots,d$, the limited slopes $\widetilde{\bcC}_{|\vect{1}|}^{i}$ by comparing the variation within $\bK$, $\bcC_{|\vect{1}|}^{i}$, with the variation in the cell average by considering the neighboring elements
\begin{equation}
  \widetilde{\bcC}_{|\vect{1}|}^{i}
  =\text{minmodB}
  \big(\,
    \bcC_{|\vect{1}|}^{i},\,
    \beta\,(\langle\bcM\rangle_{\bK}-\langle\bcM\rangle_{\bK}^{i,-}),\,
    \beta\,(\langle\bcM\rangle_{\bK}^{i,+}-\langle\bcM\rangle_{\bK})
  \,\big),
  \label{eq:limitedSlopes}
\end{equation}
where the limiter parameter $\beta\in[1,2]$ controls how aggressive limiting is applied.  
($\beta=2$ is the least aggressive, and, in many cases, leads to the most accurate results.)  
The $\text{minmodB}$ function is given by \citet{cockburnShu_2001} as
\begin{equation}
  \mbox{minmodB}(a,b,c)
  =\left\{\begin{array}{ll}
  a \quad\text{if}\quad |a|\le M\,\Delta x^{2} \\
  \mbox{minmod}(a,b,c) \quad \text{otherwise},
  \end{array}\right.
  \label{eq:minmodB}
\end{equation}
and the usual $\text{minmod}$ function is given by
\begin{equation}
  \mbox{minmod}(a,b,c)
  =\left\{\begin{array}{ll}
  s\,\min[\,|a|,|b|,|c|\,] \quad\text{if}\quad s=\mbox{sign}(a)=\mbox{sign}(b)=\mbox{sign}(c) \\
  0 \quad \text{otherwise}.
  \end{array}\right.
  \label{eq:minmod}
\end{equation}
In Eq.~\eqref{eq:minmodB}, the parameter $M>0$ is proportional to the second derivative, and can be adjusted to prevent clipping at smooth extrema.  
For $M=0$, the $\text{minmodB}$ functions reduces to the usual minmod function in Eq.~\eqref{eq:minmod}.  
In Eq.~\eqref{eq:limitedSlopes}, $\langle\bcM\rangle_{\bK}^{i,-/+}$ is the cell average in the element to the left/right in the $i$-th dimension of the element under consideration.

@@

Unfortunately, the $\text{minmod}$ function in Eq.~\eqref{eq:minmod} can result in excessive dissipation around smooth extrema and destroy high-order accuracy.  
The $\text{minmodB}$ in Eq.~\eqref{eq:minmodB} prevents this excessive dissipation, but introduces the parameter $M$, which we have found to be difficult to adjust in order to obtain uniformly acceptable results across all the numerical experiments presented later.  
To alleviate this dependence on the parameter $M$, we have implemented a discontinuity detector based on the subcell resolution idea of \citet{harten_1989}, following the description in \citet{qiuShu_2005}.

@@

As we mentioned in Section~\ref{sec:twoMomentModel} when discussing Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}, the total opacity (excluding NES) is $\kappa(\epsilonNu,\bU)=\tilde{\chi}(\epsilonNu,\bU)+\sigma(\epsilonNu,\bU)$, where the absorption opacity $\tilde{\chi}(\epsilonNu,\bU)$, corrected for stimulated absorption, is given by the sum of the neutrino emissivity $\eta(\epsilonNu,\bU)$ and the neutrino absorptivity $\chi(\epsilonNu,\bU)$
\begin{equation}
  \tilde{\chi}(\epsilonNu,\bU) = \eta(\epsilonNu,\bU) + \chi(\epsilonNu,\bU). 
\end{equation}
The scattering opacity $\sigma(\epsilonNu,\bU)$ is due to (isotropic) elastic scattering.   
Neutrino-electron scattering gives rise to the integral operators involving the rates $R^{\In}(\epsilonNu,\epsilonNu',\bU)$ and $R^{\Out}(\epsilonNu,\epsilonNu',\bU)$.  
(In the following, to simplify the notation, we will drop the explicit spatial dependence and dependence on material properties $\bU$.)  
To arrive at the collision terms for the two-moment model, we consider the space-homogeneous Boltzmann equation \citep[e.g., Eq.~(7) in][]{mezzacappaBruenn_1993c}
\begin{align}
  \f{1}{c}\deriv{f}{t}(\omegaNu,\epsilonNu,t)
  & = \eta(\epsilonNu) - \tilde{\chi}(\epsilonNu)\,f(\omegaNu,\epsilonNu,t) \nonumber \\
  & \hspace{12pt}
  + \frac{\epsilonNu^2}{c\,(hc)^3} \int_{\bbS^{2}} R_{\IS}(\bl'\cdot\bl,\epsilonNu)\,f(\omega', \epsilonNu,t)\,d\omegaNu'
  - \frac{\epsilonNu^2}{c\,(hc)^3} f(\omegaNu,\epsilonNu,t) \int_{\bbS^{2}} R_{\IS}(\bl'\cdot\bl,\epsilonNu)\,d\omega' \nonumber \\
  & \hspace{12pt} 
  + \frac{1}{c\,(hc)^3}\big(1- f(\omegaNu,\epsilonNu)\big) \, 
  \int_{\bbR^{+}} \int_{\bbS^{2}}  R^{\In}_{\NES} (\bl'\cdot\bl,\epsilonNu,\epsilonNu')\,f(\omega',\epsilonNu',t)\,d\epsilonNu' \epsilonNu'^{2}\,d\omega' \nonumber \\
  & \hspace{12pt} 
  - \frac{1}{c\,(hc)^3} f(\omegaNu,\epsilonNu,t) \int_{\bbR^{+}} 
  \int_{\bbS^{2}}  R^{\Out}_{\NES} (\bl'\cdot\bl,\epsilonNu,\epsilonNu^{\prime})\,\big(1-f(\omegaNu',\epsilonNu',t )\big)\,d\epsilonNu' \epsilonNu'^{2}\,d\omegaNu',
  \label{eq:homogeneousBoltzmann}
\end{align}
where we have brought back physical constants; i.e., the speed of light $c$ and the Planck constant $h$.  
In the scattering kernels, $\bl(\omegaNu)=(\cos\vartheta,\sin\vartheta\cos\varphi,\sin\vartheta\sin\varphi)^{\mbox{\tiny T}}$, so that
\begin{equation}
 \bl'\cdot\bl  = \mu\mu' + [(1-\mu^2)(1-\mu'^{2})]^{1/2}\cos(\phiNu-\phiNu')
 \equiv \cos\alpha,
 \quad(\mu=\cos\thetaNu)
\end{equation}
is the cosine of the angle between incoming and outgoing neutrino is the scattering process.  
In Eq.~\eqref{eq:homogeneousBoltzmann}, the first two terms are due to emission and absorption, the third and fourth terms are due to elastic (isoenergetic) neutrino scattering on nucleons and nuclei, and the last two terms are due to inelastic (non-isoenergetic) neutrino-electron scattering.

@@

The NES scattering kernel $R_{\NES}^{\Out}(\cos\alpha,\epsilonNu,\epsilonNu')$ is given by Eq.~(9) in \citet{mezzacappaBruenn_1993c}.  
As we did in the isoenergetic scattering case, we approximate the angular dependence of the scattering kernels by a two-term Legendre expansion; e.g.,
\begin{equation}
  R^{\Out}_{\NES}(\cos\alpha,\epsilonNu,\epsilonNu') 
  \approx \f{1}{2} \Phi_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu') + \f{3}{2}\Phi^{\Out}_{\NES,1}(\epsilonNu,\epsilonNu')\,\cos\alpha,
\end{equation}
where
\begin{equation}
  \Phi_{\NES,\ell}^{\Out}(\epsilonNu,\epsilonNu')
  =\int_{-1}^{+1}R_{\NES}^{\Out}(\cos\alpha,\epsilonNu,\epsilonNu')\,P_{\ell}(\cos\alpha)\,d\cos\alpha \quad (\ell=0,1).  
\end{equation}
(Again, these integrals are computed numerically using a ??-point Gauss-Legendre quadrature.  (\ee{Ran: Fill in the number of quadrature points}.))
The in-scattering kernels $\Phi_{\NES,\ell}^{\In}(\epsilonNu,\epsilonNu')$ are obtained from symmetry properties \citep{cernohorsky_1994}
\begin{equation}
  \Phi_{\NES,\ell}^{\In}(\epsilonNu,\epsilonNu')=\Phi_{\NES,\ell}^{\Out}(\epsilonNu',\epsilonNu)
  \quad\text{and}\quad
  \Phi_{\NES,\ell}^{\Out}(\epsilonNu',\epsilonNu)
  =e^{-(\epsilonNu-\epsilonNu')/k_{\mbox{\tiny B}}T}\,\Phi_{\NES,\ell}^{\Out}(\epsilonNu,\epsilonNu').
  \label{eq:symmetryNES}
\end{equation}
Inserting the Legendre expansion into Eq.~\eqref{eq:homogeneousBoltzmann}, and taking the zeroth moment (number equation), gives
\begin{align}
  \f{1}{c}\deriv{\cJ}{t}
  &=\big(\,4\pi-\cJ(\epsilonNu)\,\big)\int_{\bbR^{+}}R_{\NES,0}^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  -\bcH(\epsilonNu)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\In}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'} \nonumber \\
  & \hspace{12pt}
  -\cJ(\epsilonNu)\int_{\bbR^{+}}R_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu')\,\big(\,4\pi-\cJ(\epsilonNu')\,\big)\,dV_{\epsilonNu'}
  +\bcH(\epsilonNu)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\Out}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'}.  
\end{align}
Similarly, taking the first moment (number flux) of Eq.~\eqref{eq:homogeneousBoltzmann} results in
\begin{align}
  \f{1}{c}\deriv{\bcH}{t}
  &=-\bcH(\epsilonNu)\int_{\bbR^{+}}R_{\NES,0}^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  +\big(\,(4\pi/3)\,\bI-\bcK\,\big)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\In}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'} \nonumber \\
  & \hspace{12pt}
  -\bcH\int_{\bbR^{+}}R_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu')\,\big(\,4\pi-\cJ(\epsilonNu')\,\big)\,dV_{\epsilonNu'}
  +\bcK\cdot\int_{\bbR^{+}}R_{\NES,1}^{\Out}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'},
\end{align}
where we have defined
\begin{equation}
  R_{\NES,0}^{\In/\Out}(\epsilonNu,\epsilonNu')
  =\f{1}{c\,(hc)^{3}}\,\f{1}{2}\,\Phi_{\NES,0}^{\In/\Out}(\epsilonNu,\epsilonNu')
  \quad\text{and}\quad
  R_{\NES,1}^{\In/\Out}(\epsilonNu,\epsilonNu')
  =\f{1}{c\,(hc)^{3}}\,\f{3}{2}\,\Phi_{\NES,1}^{\In/\Out}(\epsilonNu,\epsilonNu').  
\end{equation}
We tabulate the scattering kernels $R_{\NES,0}^{\Out}$ and $R_{\NES,1}^{\Out}$, which have units MeV$^{-3}$ cm$^{-1}$, and obtain the in-scattering rates from the symmetry properties in Eq.~\eqref{eq:symmetryNES}.  
We only use $R_{\NES,0}^{\In/\Out}$ in the numerical experiments in Section~\ref{sec:numericalTableOpacities}, so that $R^{\In/\Out}=R_{\NES,0}^{\In/\Out}$ in Eqs~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}.

@@

The line source benchmark \citep[cf.][]{brunner_2002,garrettHauck_2013} is a challenging test for approximate transport algorithms.  
In cylindrical coordinates, it consists of an initial delta function particle distribution in radius; i.e., $f_{0}=\delta(R)/4\,\pi$.  
For $t>0$, a cylindrical radiation front propagates radially.  
Apart from capturing details of the exact transport solution, maintaining realizability ($\bcM\in\cR$ cf. Eq.~\eqref{eq:realizableSet}) of the two-moment solution is challenging.

This test is computed using cylindrical coordinates on a one-dimensional domain with $R\in[0,1.5]$.  
To avoid initiating the radiation field with the delta function, we follow the procedure in \citet{garrettHauck_2013}, and approximate the initial condition with an isotropic Gaussian distribution function
\begin{equation}
  f_{\mbox{\tiny G},0}
  =\max\Big[\,\f{1}{8\,\pi\,\sigma_{\mbox{\tiny G}}^{2}}\,e^{-R^{2}/(2\,\sigma_{\mbox{\tiny G}}^{2})},10^{-4}\,\Big],
\end{equation}
where we use $\sigma_{\mbox{\tiny G}}=0.03$.  
We run this test to a final time $t=1.0$.

In Figure~\ref{fig:lineSource} we plot the density $\cJ$ (left panel) and flux factor $h=|\cH_{1}|/\cJ$ (right panel) for various times during the evolution: $t=0.0$ (red), $t=0.35$ (green), $t=0.65$ (blue), and $t=1.0$ (black).  
The solid lines are for a high-resolution reference solution obtained with the DG(1)+RK2 scheme using 2048 elements, while the open circles are for a low resolution run using the DG(2)+RK3 scheme and 64 elements.  
There is good agreement between the two solutions.  
When comparing the density $\cJ$ with the transport solutions in \citet{garrettHauck_2013}, we observe that the peak of the radiation front reaches the correct location, but otherwise, the two-moment model is not accurate for this problem.  
This is expected since a high number of angular moments is needed to capture the transport solution.  
Despite the shortcomings of the two-moment approximation, our implementation maintains $\cJ\ge0$ and $h\in[0,1]$.

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \includegraphics[scale=0.4]{./Figures/LineSource1D_Density.png} &
      \includegraphics[scale=0.4]{./Figures/LineSource1D_FluxFactor.png}
    \end{tabular}
  \end{center}
  \caption{Results from the line source problem with the two-moment model for various times.  In the left panel we plot the radiation density $\cJ$.  In the right panel we plot the flux factor $h=|\cH_{1}|/\cJ$.  Solid lines represent a high resolution (2048 elements) reference solution using the DG(1)+RK2 scheme, while circles are results obtained with the DG(2)+RK3 scheme using 64 elements.  Both solutions maintain the bounds $\cJ>0$ and $h\in[0,1]$.}
  \label{fig:lineSource}
\end{figure}
